Using Metric Space Indexing for Complete and Efficient Record Linkage

### 2017-11-21 paper.tex
Update on Overleaf.

\emph{Metric space indexing} (MSI) is a complete technique with much
lower computational overheads than a brute-force approach. It allows
indexing, comparison and classification to be combined into a single
step, as shown in Fig.~\ref{fig-rl-process}, making the overall record
linkage process simpler, more efficient and more effective than
incomplete indexing approaches.

@@

We address the following general record linkage problem: for two
datasets $A$ and $B$, we wish to find, for each record in $A$, all the
records in $B$ that match it. We compare several algorithms for record
linkage: traditional blocking, one incomplete similarity search method,
LSH-MinHash, and one complete method, M-tree~\cite{paolociaccia2m}. We
also use a simple complete brute-force technique as a baseline for
comparison, although this can only feasibly be applied to the smallest
of our datasets. All experiments have a number of parameters to
configure the search space and algorithm behaviour, including a
\emph{distance threshold} specifying a maximum distance, equivalent to a
minimum similarity, for two records to be classified as a link (i.e.\
referring to the same entity).

\textbf{Brute-force:}
Two nested loops are used to compare every record in $A$ with every
record in $B$. Each pair is classified as a link if the distance between
the records is less than the threshold. This approach is guaranteed to
identify all links, with complexity $ O(|A| \times |B|) $.

@@

\Cref{cora-quality} shows the precision, recall, and F-measure
for various thresholds. As expected, low thresholds give high precision and low recall, and the reverse for high threshold
values. As expected, brute force and

For a more detailed investigation of selected linkers, the brute-force
approach is used to establish the optimum threshold value for the Cora
dataset. The maximum F-measure value is observed at a threshold value of
70. This value is dataset-dependent; for different datasets the maximum
F-measure will occur at different thresholds. In the rest of this
section we fix the threshold value at 70.

@@

M-tree yields the same linkage quality as brute-force, although using a
significantly lower number of comparisons. This is as expected, since
both techniques are complete. Several of the incomplete linkers give
similar quality, for example \emph{LSH-2-2}, \emph{LSH-5-2}, \emph{LSH-10-2} and
\emph{Block-combined}. These, and a number of other incomplete linkers give
better precision than the complete techniques. This is due to
high-similarity non-matches, as discussed in Sect.~\ref{sec-approach}.
Although several of the incomplete linkers give as good quality as
M-tree, and in some cases at lower cost, this is offset by the need to
select appropriate configuration parameters. Some other linkers give
very poor results.

@@

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
\caption{Cora - Brute Force}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
\caption{Cora - M-tree}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
\caption{Cora - Traditional Blocking on `authors' field}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
\caption{Cora - Traditional Blocking on `title' field}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
\caption{Cora - Traditional Blocking on `publication year' field}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
\caption{Cora - Traditional Blocking using the union of blocking on all fields}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-2}
\caption{Cora - LSH Blocking using 2 bands of size 2}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-10}
\caption{Cora - LSH Blocking using 2 bands of size 10}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-2}
\caption{Cora - LSH Blocking using 10 bands of size 2}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-10}
\caption{Cora - LSH Blocking using 10 bands of size 10}
\end{subfigure}

\caption{Linkage results on Cora dataset}
\label{cora-quality}
\end{figure}

@@

\Cref{comparison-of-results-demography-skye} and \Cref{comparison-of-results-demography-kili} show the results for LSH-Minhash and M-tree on the Skye and Kilmarnock datasets. Linkage was performed between birth and death records. The shingle size was set to 2 for all the LSH experiments reported, as this was found to give good results and LSH was not especially sensitive to this parameter. Results for other shingle sizes were omitted from this paper for brevity. In both cases the F1 measure achieved is better for M-tree than any of the experimental settings chosen for LSH. This is largely due to the recall for M-tree being much higher than any of the LSH experiments. For all LSH parameters LSH out performs M-tree in terms of precision.

@@

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/skye-mtree}
\caption{Linkage quality on Skye dataset using M-tree\label{skye-quality-mtree}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/kili-mtree}
\caption{Linkage quality on Kilmarnock dataset using M-tree\label{kilmarnock-quality-mtree}}
\end{subfigure}

@@

In this paper we have demonstrated the efficacy of MSI in achieving  complete and efficient record linkage, without the need for complex parameter tuning. In conclusion, this assertion deserves some careful unpacking. It is always possible to achieve high quality linkage using a brute-force approach. However the n-squared complexity of this approach prevents its practical application for datasets of even moderate size. The use of MSI techniques such as M-tree delivers the same results as the brute-force approach with fewer comparisons. We have shown that the M-tree approach delivers high precision, high recall results that are the same as those delivered by the brute-force approach. Furthermore this is achieved consistently without the need for complex parameter tuning.

@@

However, this paper also highlights some unexpected results. Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute-force as the number of records in the block increase. A second unexpected result is that shown in \Cref{comparison-of-results-cora}, namely that LSH can yield higher precision than is achieved by a complete method such as M-tree, as was discussed in Sect.~\ref{sec-approach}.

>>>>

\emph{Metric space indexing} (MSI) is a complete technique with much
lower computational overheads than a brute force approach. It allows
indexing, comparison and classification to be combined into a single
step, as shown in Fig.~\ref{fig-rl-process}, making the overall record
linkage process simpler, more efficient and more effective than
incomplete indexing approaches.

@@

We address the following general record linkage problem: for two
datasets $A$ and $B$, we wish to find, for each record in $A$, all the
records in $B$ that match it. We compare several algorithms for record
linkage: traditional blocking, one incomplete similarity search method,
LSH-MinHash, and one complete method, M-tree~\cite{paolociaccia2m}. We
also use a simple complete brute force technique as a baseline for
comparison, although this can only feasibly be applied to the smallest
of our datasets. All experiments have a number of parameters to
configure the search space and algorithm behaviour, including a
\emph{distance threshold} specifying a maximum distance, equivalent to a
minimum similarity, for two records to be classified as a link (i.e.\
referring to the same entity).

\textbf{Brute force:}
Two nested loops are used to compare every record in $A$ with every
record in $B$. Each pair is classified as a link if the distance between
the records is less than the threshold. This approach is guaranteed to
identify all links, with complexity $ O(|A| \times |B|) $.

@@

\Cref{cora-quality} shows the precision, recall, and F-measure
for various thresholds. As expected, low thresholds give high precision and low recall, and the reverse for high threshold
values. As expected, brute force and M-tree give identical results.

For a more detailed investigation of selected linkers, the brute force
approach is used to establish the optimum threshold value for the Cora
dataset. The maximum F-measure value is observed at a threshold value of
70. This value is dataset-dependent; for different datasets the maximum
F-measure will occur at different thresholds. In the rest of this
section we fix the threshold value at 70.

@@

M-tree yields the same linkage quality as brute force, although using a
significantly lower number of comparisons. This is as expected, since
both techniques are complete. Several of the incomplete linkers give
similar quality, for example \emph{LSH-2-2}, \emph{LSH-5-2}, \emph{LSH-10-2} and
\emph{Block-combined}. These, and a number of other incomplete linkers give
better precision than the complete techniques. This is due to
high-similarity non-matches, as discussed in Sect.~\ref{sec-approach}.
Although several of the incomplete linkers give as good quality as
M-tree, and in some cases at lower cost, this is offset by the need to
select appropriate configuration parameters. Some other linkers give
very poor results.

@@

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
\caption{Brute Force}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
\caption{M-tree}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
\caption{Blocking on `authors' field}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
\caption{Blocking on `title' field}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
\caption{Blocking on `publication year' field}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
\caption{Blocking using the union of all fields}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-2}
\caption{LSH using 2 bands of size 2}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-10}
\caption{LSH using 2 bands of size 10}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-2}
\caption{LSH using 10 bands of size 2}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-10}
\caption{LSH using 10 bands of size 10}
\end{subfigure}

\caption{Linkage results on the Cora dataset}
\label{cora-quality}
\end{figure}

@@

\Cref{comparison-of-results-demography-skye} and \Cref{comparison-of-results-demography-kili} show the results for LSH-MinHash and M-tree on the Skye and Kilmarnock datasets. Linkage was performed between birth and death records. The shingle size was set to 2 for all the LSH experiments reported, as this was found to give good results and LSH was not especially sensitive to this parameter. Results for other shingle sizes were omitted from this paper for brevity. In both cases the F1 measure achieved is better for M-tree than any of the experimental settings chosen for LSH. This is largely due to the recall for M-tree being much higher than any of the LSH experiments. For most configurations, LSH out-performs M-tree in terms of precision.

@@

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
\caption{Linkage quality on Skye dataset using M-tree\label{skye-quality-mtree}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
\caption{Linkage quality on Kilmarnock dataset using M-tree\label{kilmarnock-quality-mtree}}
\end{subfigure}

@@

In this paper we have demonstrated the efficacy of MSI in achieving  complete and efficient record linkage, without the need for complex parameter tuning. In conclusion, this assertion deserves some careful unpacking. It is always possible to achieve high quality linkage using a brute force approach. However the n-squared complexity of this approach prevents its practical application for datasets of even moderate size. The use of MSI techniques such as M-tree delivers the same results as brute force with fewer comparisons. We have shown that the M-tree approach delivers high precision, high recall results that are the same as those delivered by brute force. Furthermore this is achieved consistently without the need for complex parameter tuning.

@@

However, this paper also highlights some unexpected results. Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute force as the number of records in the block increase. A second unexpected result is that shown in \Cref{comparison-of-results-cora}, namely that LSH can yield higher precision than is achieved by a complete method such as M-tree, as was discussed in Sect.~\ref{sec-approach}.
